---
pdf_document: default
author: "Enzo Ramirez, Pauline De taeye"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
html_document:
  df_print: paged
title: "Bootstrap Lab"
word_document: default
---
  
```{r Knitr_Global_Options, include=FALSE, cache=TRUE}
setwd('V:/export')

# Donner le nom du répertoire pour installer les librairies
.libPaths("K:/M2ST/Lib")
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, 
               autodep = TRUE, tidy = FALSE,
               fig.dim=c(8,3))
```


```{r Libraries, cache=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
library(boot)
#install.packages("foreach")
library(foreach)
library(pracma)
```


# Ex 2 
  
Now redo Ex 1 Questions 1, 2, 4 to 7 focusing on estimating the variance of $x$. 
  

In this exercise, we use the package **boot** to calculate a bootstrap estimate of the variance of the sample mean and build a confidence interval. 

  1. Using the function *rchisq*, generate a random sample of size $n=25$ from the $\chi^2_1$ distribution. Denoting the sample variance estimator as $V(x)$, what is the value of $V(x)$ on this particular sample? Compare with the population variance.


```{r}
set.seed(2020)
n <- 25
mysample <- rchisq(n, df = 1)
Population.var <- 2
cat("\n", "sample variance : ", var(mysample), "\n", "population variance : "  , Population.var)
```

We can see that the mean estimate is larger than the population


  2. Compute theoretically the variance and standard deviation of the estimator $V(x)$ in the present setup, that is, for a random sample of size $25$ from the $\chi^2_1$ distribution?


```{r}
mean(mysample)
std(mysample)
var(mysample)
```



    4.


```{r}
boot.var <- function(asample, indices) {
  var(asample[indices])
}
```

  5. Use the *boot* function to generate $B=10000$ sample var of bootstrap samples. Store the results in an object.
Which information is included in this object?
Compute the resulting bootstrap estimate of the variance of $\bar{x}$. 


```{r}
B <- 10000
# tic()
res.boot <- boot(mysample,boot.var,B)
# toc()
names(res.boot)
```



  6. What are the sample mean and the sample variance of the bootstrap variance?



```{r}
mean(res.boot$t)
var(res.boot$t)
sd(res.boot$t)
```



  7. Show a histogram of the $B=10000$ sample means. Identify the population mean, the sample mean of the original sample, and the mean of the bootstrap means.



```{r}
hist(res.boot$t, col="lightgrey", xlab="Sample Variance", breaks=50,   main="Distribution of the Variance of B=10000 bootstrap samples")
Population.mean <- 1
Population.var <- 2
abline(v=Population.var, lwd=2, col="blue")
abline(v=var(mysample), lwd=2, col="red")
abline(v=mean(res.boot$t), lwd=2, col="orange")
legend("topleft", lwd=2, col=c("blue", "red", "orange"), legend=c("Population variance", "Sample mean", "Mean of bootstrap Variance"))
```



# Ex 3

This exercise illustrates how the bootstrap can be used in linear regression to obtain (i) confidence intervals for the parameters and (ii) confidence bounds for the prediction.

For $\alpha=-1$ and $\beta=2$, generate a random sample $(X_1,Y_1),\ldots,(X_{20},Y_{20})$ from the linear regression model :
    
$Y = \alpha + \beta X+ 0.5 (\varepsilon-1)$
  

Where $X$ is uniformly distributed over $[0,1]$,

$\varepsilon$ is chi-square distributed  with 1 degree of freedom,

and where $X$ and $\varepsilon$ are independent

(note that, in this model, the least squares estimators of $\alpha$ and $\beta$ *do not* follow the usual normal distribution, which makes inference more challenging). 


```{r}
set.seed(2020)
n <- 20
alpha <- -1
beta <- 2

epsilon <- rchisq(n, df = 1)
mean(epsilon)

X <- runif(n, 0, 1)
mean(X)

Y = alpha +beta*X + 0.5 * (epsilon -1)

df <- data.frame("Y" = Y, "X" = X)
```



  1. Make a scatter plot of this sample. Add to the plot the true linear relationship, namely the straight line $y=2x-1$. 



```{r}
plot(X, Y, main="Scatterplot",
    xlab="X var",
    ylab="Y var",
    pch=19)
# abline(lm(Y ~ X), col = "blue")
lines(lowess(X,2*X-1))
```



  2. What are the least-squares estimators? Add to the plot the corresponding estimated linear regression line.



```{r}
plot(X, Y, main="Scatterplot",
    xlab="X var",
    ylab="Y var",
    pch=19)
abline(lm(Y ~ X), col = "blue")
lines(lowess(X,2*X-1))
```




  3. Use the function *boot* to find a 95\% confidence interval for $\hat{\alpha}$ from $B=5000$ bootstrap samples. Repeat the exercise for $\hat{\beta}$.



```{r}
B <- 5000

paste("------------------------------------------------------------------------")

# tic()
boot.reg <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=d)
  return(coef(fit))
}

# bootstrapping with B replications
x.boot <- boot(data=df,
               statistic=boot.reg,
               R=B,
               formula=Y~X)
# toc()

names(x.boot)
x.coef.o <- as.data.frame(t(x.boot$t0))

paste("------------------------------------------------------------------------")

x.boot

paste("------------------------------------------------------------------------")

# colMeans(x.boot$t)

cat( "mean alpha : ", mean(x.boot$t[,1]),  "/ mean beta : ", mean(x.boot$t[,2] ))
cat("\n")
cat( "variance alpha : ", var(x.boot$t[,1]), "/ variance beta : ", var(x.boot$t[,2] ))


paste("------------------------------------------------------------------------")

q1.alpha <- quantile(x.boot$t[,1] - x.coef.o[,1], probs = 0.025)
q2.alpha <- quantile(x.boot$t[,1] - x.coef.o[,1], probs = 0.975)

q1.beta <- quantile(x.boot$t[,2] - x.coef.o[,2], probs = 0.025)
q2.beta <- quantile(x.boot$t[,2] - x.coef.o[,2], probs = 0.975)

cat("alpha est compris entre : ", c(x.coef.o[,1] - q2.alpha, x.coef.o[,1] - q1.alpha))
cat("\n")
cat("beta est compris entre : " , c(x.coef.o[,2] - q2.beta, x.coef.o[,2] - q1.beta))

paste("------------------------------------------------------------------------")

boot.ci(x.boot)

paste("------------------------------------------------------------------------")

```



  4. Use the function *boot* to obtain a 95\% confidence interval for each of the predictions associated with $x=0, 0.05, 0.1, \ldots,0.95, 1$ (use $B=5000$ bootstrap samples).\
  Add the corresponding lower and upper bounds for the prediction (as functions of $x$) to the previous plot.



```{r}
x.boot <- boot(data=df,
                   statistic=boot.reg,
                   R=B,
                   formula=Y~X)
```





```{r}
B <- 5000

boot.reg <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample
  fit <- lm(formula, data=d)
  return(coef(fit))
}

vec.down <- c()
vec.up <- c()
vec.x <- c()


paste("------------------------------------------------------------------------")
for (xx in seq(0.05, 1, by=0.05)){
    print(paste("The x is", xx))
    x <- rep(xx, n)

    Y = alpha +beta*x + 0.5 * (epsilon -1)

    df1 <- data.frame("Y" = Y, "X" = X)

    paste("------------------------------------------------------------------------")

    # tic()

    # bootstrapping with B replications
    x.boot <- boot(data=df1,
                   statistic=boot.reg,
                   R=B,
                   formula=Y~X)
    # toc()

    names(x.boot)
    x.coef.o <- as.data.frame(t(x.boot$t0))

    paste("------------------------------------------------------------------------")

    x.boot

    paste("------------------------------------------------------------------------")

    # colMeans(x.boot$t)

    cat( "mean alpha : ", mean(x.boot$t[,1]),  "/ mean beta : ", mean(x.boot$t[,2] ))
    cat("\n")
    cat( "variance alpha : ", var(x.boot$t[,1]), "/ variance beta : ", var(x.boot$t[,2] ))
    cat("\n")
    cat("\n")


    paste("------------------------------------------------------------------------")

    q1.alpha <- quantile(x.boot$t[,1] - x.coef.o[,1], probs = 0.025)
    q2.alpha <- quantile(x.boot$t[,1] - x.coef.o[,1], probs = 0.975)

    q1.beta <- quantile(x.boot$t[,2] - x.coef.o[,2], probs = 0.025)
    q2.beta <- quantile(x.boot$t[,2] - x.coef.o[,2], probs = 0.975)

    cat("alpha est compris entre : ", c(x.coef.o[,1] - q2.alpha, x.coef.o[,1] - q1.alpha))
    cat("\n")
    cat("beta est compris entre : " , c(x.coef.o[,2] - q2.beta, x.coef.o[,2] - q1.beta))
    cat("\n")
    
    y.down <- (x.coef.o[,1] - q2.alpha) + (x.coef.o[,2] - q2.beta)*xx
    y.up <- (x.coef.o[,1] - q1.alpha) + (x.coef.o[,2] - q1.beta)*xx
    
    cat( "y est compris entre : " , c( y.down  , y.up ))
    
    vec.down <- append(vec.down, y.down)
    vec.up <- append(vec.up, y.up)
    vec.x <- append(vec.x, xx)

    cat("\n")
    cat("\n")
    cat("\n")
    

    
    
}

plot.conf <- data.frame("x" = vec.x,
                          "down" = vec.down,
                          "upper" = vec.up)
df$upper <- vec.up
df$down <- vec.down
df$xx <- vec.x

plot(df$X, df$Y, main="Scatterplot",
    xlab="X var",
    ylab="Y var",
    pch=20)
abline(lm(df$Y ~ df$X), col = "blue")
lines(lowess(X,2*X-1))
points(df$xx, df$down, col = "red", pch = 20)
points(df$xx, df$upper, col = "blue", pch = 20)


paste("------------------------------------------------------------------------")
```
